Portfolios of Subgraph Isomorphism Algorithms

### 2015-12-11 paper/paper.tex
Patrick doesn't love us any more

\author{
    Lars Kotthoff\inst{1}
    \and Ciaran McCreesh\thanks{This work was supported by the Engineering
        and Physical Sciences Research Council [grant number EP/K503058/1]}\inst{2}
    \and Patrick Prosser\inst{2}
    \and Christine Solnon\inst{3}}

@@

\paragraph{Filtering for subgraph isomorphism.}

@@

\paragraph{Propagation of invariant properties.}

@@

\paragraph{Algorithms selected in our portfolio.}
We have selected a set of algorithms with complementary performance:
\begin{itemize}
\item VF2 \cite{Cordella:2004}, which performs a weak filtering so that it is especially fast on trivially satisfiable instances;
\item LAD \cite{Solnon:2010}, which combines two strong but expensive filtering techniques (AllDifferent Generalized Arc Consistency and LAD);
\item Glasgow \cite{McCreesh:2015}, which does expensive preprocessing based on path length
    invariant properties to generate additional constraints, followed by weaker filtering
    (forward-checking, and a heuristic AllDifferent propagator which can miss deletions) and
    conflict-directed backjumping during search.
\end{itemize}

@@

\section{Experimental Comparison of Algorithms of the Portfolio}\label{expComp}

\paragraph{Benchmark.}

Note that Classes 3 and 12 were not considered in the previous experimental study reported in \cite{McCreesh:2015}.

\paragraph{Experimental setup.} We measured runtimes on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running
Scientific Linux 6.5. We used the C++ implementation of the Glasgow algorithm \cite{McCreesh:2015},
the C implementation of LAD \cite{Solnon:2010}, and the VFLib C implementation of VF2
\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
timeout of $10^8$ milliseconds (a little over a day).

\begin{figure}[t]
\includegraphics[width=\textwidth]{courbe.pdf}
\caption{Evolution of the number of solved instances with respect to CPU time, for the 8 algorithms of the portfolio, and a Virtual Best Solver (VBS).\label{expTime}}
\end{figure}

\paragraph{Results.}Figure \ref{expTime} displays the evolution of the number of instances solved with respect to CPU time. It shows us that the best solver depends on the time limit
considered. IncompleteLAD is able to solve easy inconsistent instances very quickly, in a few milliseconds. Hence, for time limits lower than 5 ms, the best solver is incompleteLAD. However, it is not able to solve harder inconsistent instances, neither can it solve consistent instances.

@@

\medskip

Battiti and Mascia~\cite{battiti-mascia07} were the first to propose
algorithm portfolios for subgraph isomorphism problems. However, they only
consider a pure parallel portfolio without a selection mechanism consisting of
only two solvers. In addition, our instance set is much larger and considers
both hard and unsatisfiable instances.

@@

\begin{table}[t]
\begin{center}
\begin{tabular}{|c||r||r|r|r||r|r|r|r||r|r|}
\hline
$t$ & VF2 & \multicolumn{3}{c||}{LAD} & \multicolumn{4}{c||}{Glasgow}& VBS & $\delta$\\
&&pre&def&new&1&2&3&4&&\\\hline
$10^0$ & 300 &  \cellcolor{blue!25}1126 & 661 & 599 & 386 & 95 & 39 & 12 & 1132 & 6\\\hline
$10^1$ & 1699 & 1858 & 2101 &  \cellcolor{blue!25}2287 & 2061 & 1215 & 455 & 208 & 3388 & 101\\\hline
$10^2$ & 2834 & 1912 & 3376 & 3708 &  \cellcolor{blue!25}3943 & 3344 & 2295 & 1329 & 4636 & 693\\\hline
$10^3$ & 3458 & 1918 & 4233 & 4531 &  \cellcolor{blue!25}4908 & 4763 & 4262 & 3426 & 5170 & 262\\\hline
$10^4$ & 3701 & 1919 & 4888 & 5025 & 5156 &  \cellcolor{blue!25}5253 & 5022 & 4408 & 5332 & 79\\\hline
$10^5$ & 3850 & 1919 & 5108 & 5193 & 5301 &  \cellcolor{blue!25}5377 & 5292 & 5082 & 5434 & 57\\\hline
$10^6$ & 3977 & 1919 & 5244 & 5307 & 5386 &  \cellcolor{blue!25}5452 & 5451 & 5235 & 5500 & 48\\\hline
$10^7$ & 4082 & 1919 & 5336 & 5414 & 5458 & 5517 &  \cellcolor{blue!25}5518 & 5425 & 5567 & 49\\\hline
$10^8$ & 4191 & 1919 & 5423 & 5479 & 5508 &  \cellcolor{blue!25}5561 & 5560 & 5554 & 5608 & 47\\\hline
\end{tabular}
\end{center}
\caption{Number of solved instances at different CPU time limits: Each line displays a time limit
$t$ (in milliseconds) followed by the number of instances solved within this time limit by VF2, LAD
preprocessing, default LAD, and new LAD, Glasgow with  the lengths of paths limited to 1, 2, 3 and
4, and the VBS. The best single algorithm is highlighted in blue, and the difference between VBS and
single best if displayed in the last column ($\delta$).\label{expTime}}
\end{table}

Table \ref{expTime} displays the number of instances solved at different CPU time limits, ranging
from 1 to $10^8$ milliseconds. It shows us that the best single solver depends on the time limit
considered. Simple approaches like LAD preprocessing and VF2 are able to solve easy instances very
quickly, in a few milliseconds. However, they are not able to solve harder instances: LAD
preprocessing is an incomplete approach which can only detect rather trivial unconsistencies; VF2
performs a basic backtracking search and is very fast on easy instances, because it does not compute
expensive invariants. Default LAD, new LAD, and the Glasgow algorithms solve more instances than VF2
and LAD preprocessing when considering longer time limits: 10 ms for default and new LAD and Glasgow
1, and 100 (resp. 1000 and 10000) ms for Glasgow 2 (resp. 3 and 4). The Virtual Best Solver (VBS),
which considers the best algorithm for each instance separately, obtains much better results,
showing us that the algorithms have complementary performance. The difference between VBS and single
best (column $\delta$ of Table \ref{expTime}) is more particularly important for rather small CPU
time limits. In particular, VBS solves 693 more instances than the single best (Glasgow 1) when the
limit is 100 ms. In many applications, it is important to have the fastest possible algorithm. For
example, in pattern recognition applications, we often have to solve subgraph isomorphism problems
for a very large number of graphs (in order to find a pattern image in a large database of target
images, for example), so that having an algorithm that is able to solve an instance in 100 ms
instead of 1000 ms makes a big difference. Therefore, it is important to select the best algorithm
for each instance.

>>>>

\author{
    Lars Kotthoff\inst{1}
    \and Ciaran McCreesh\thanks{This work was supported by the Engineering
        and Physical Sciences Research Council [grant number EP/K503058/1]}\inst{2}
    \and Christine Solnon\inst{3}}

@@

\subsection{Filtering for subgraph isomorphism}

@@

\subsection{Propagation of invariant properties}

@@

\subsection{Problem Instances}

Note that Classes 3 and 12 were not considered in the previous experimental
study reported in \cite{McCreesh:2015}.  This set of instances is also much
larger than that of Battiti and Mascia~\cite{battiti-mascia07}, who were the
first to propose algorithm portfolios for subgraph isomorphism problems.
Battiti and Mascia only considered a pure parallel portfolio consisting of two
solvers, and without a selection mechanism. Their problem set consisted
entirely of easy satisfiable instances, which is where this style of approach
has most success.

\subsection{Experimental setup}

We measured runtimes on machines with Intel Xeon E5-2640 v2 CPUs and 64GBytes RAM, running
Scientific Linux 6.5. We used the C++ implementation of the Glasgow algorithm \cite{McCreesh:2015},
the C implementation of LAD \cite{Solnon:2010}, and the VFLib C implementation of VF2
\cite{Cordella:2004}. Software was compiled using GCC 4.9. Each problem instance was run with a
timeout of $10^8$ milliseconds (a little over a day).

\subsection{Algorithms selected in our portfolio}

We have selected a set of algorithms with complementary performance:
\begin{itemize}
\item VF2 \cite{Cordella:2004}, which performs a weak filtering so that it is especially fast on trivially satisfiable instances;
\item LAD \cite{Solnon:2010}, which combines two strong but expensive filtering techniques (AllDifferent Generalized Arc Consistency and LAD);
\item Glasgow \cite{McCreesh:2015}, which does expensive preprocessing based on path length
    invariant properties to generate additional constraints, followed by weaker filtering
    (forward-checking, and a heuristic AllDifferent propagator which can miss deletions) and
    conflict-directed backjumping during search.
\end{itemize}

@@

\subsection{Results from Individual Algorithms} \label{expComp}

\begin{figure}[t]
\includegraphics[width=\textwidth]{courbe.pdf}
\caption{Evolution of the number of solved instances with respect to CPU time, for the 8 algorithms of the portfolio, and a Virtual Best Solver (VBS).\label{expTimeGraph}}
\end{figure}

Figure \ref{expTimeGraph} displays the evolution of the number of instances solved with respect to CPU time. It shows us that the best solver depends on the time limit
considered. IncompleteLAD is able to solve easy inconsistent instances very quickly, in a few milliseconds. Hence, for time limits lower than 5 ms, the best solver is incompleteLAD. However, it is not able to solve harder inconsistent instances, neither can it solve consistent instances.

@@

@@

\begin{table}[t]
\begin{center}
\begin{tabular}{|c||r||r|r|r||r|r|r|r||r|r|}
\hline
$t$ & VF2 & \multicolumn{3}{c||}{LAD} & \multicolumn{4}{c||}{Glasgow}& VBS & $\delta$\\
&&pre&def&new&1&2&3&4&&\\\hline
$10^0$ & 300 &  \cellcolor{blue!25}1126 & 661 & 599 & 386 & 95 & 39 & 12 & 1132 & 6\\\hline
$10^1$ & 1699 & 1858 & 2101 &  \cellcolor{blue!25}2287 & 2061 & 1215 & 455 & 208 & 3388 & 101\\\hline
$10^2$ & 2834 & 1912 & 3376 & 3708 &  \cellcolor{blue!25}3943 & 3344 & 2295 & 1329 & 4636 & 693\\\hline
$10^3$ & 3458 & 1918 & 4233 & 4531 &  \cellcolor{blue!25}4908 & 4763 & 4262 & 3426 & 5170 & 262\\\hline
$10^4$ & 3701 & 1919 & 4888 & 5025 & 5156 &  \cellcolor{blue!25}5253 & 5022 & 4408 & 5332 & 79\\\hline
$10^5$ & 3850 & 1919 & 5108 & 5193 & 5301 &  \cellcolor{blue!25}5377 & 5292 & 5082 & 5434 & 57\\\hline
$10^6$ & 3977 & 1919 & 5244 & 5307 & 5386 &  \cellcolor{blue!25}5452 & 5451 & 5235 & 5500 & 48\\\hline
$10^7$ & 4082 & 1919 & 5336 & 5414 & 5458 & 5517 &  \cellcolor{blue!25}5518 & 5425 & 5567 & 49\\\hline
$10^8$ & 4191 & 1919 & 5423 & 5479 & 5508 &  \cellcolor{blue!25}5561 & 5560 & 5554 & 5608 & 47\\\hline
\end{tabular}
\end{center}
\caption{Number of solved instances at different CPU time limits: Each line displays a time limit
$t$ (in milliseconds) followed by the number of instances solved within this time limit by VF2, LAD
preprocessing, default LAD, and new LAD, Glasgow with  the lengths of paths limited to 1, 2, 3 and
4, and the VBS. The best single algorithm is highlighted in blue, and the difference between VBS and
single best if displayed in the last column ($\delta$).\label{expTimeTable}}
\end{table}

Table \ref{expTimeTable} displays the number of instances solved at different CPU time limits, ranging
from 1 to $10^8$ milliseconds. It shows us that the best single solver depends on the time limit
considered. Simple approaches like LAD preprocessing and VF2 are able to solve easy instances very
quickly, in a few milliseconds. However, they are not able to solve harder instances: LAD
preprocessing is an incomplete approach which can only detect rather trivial unconsistencies; VF2
performs a basic backtracking search and is very fast on easy instances, because it does not compute
expensive invariants. Default LAD, new LAD, and the Glasgow algorithms solve more instances than VF2
and LAD preprocessing when considering longer time limits: 10 ms for default and new LAD and Glasgow
1, and 100 (resp. 1000 and 10000) ms for Glasgow 2 (resp. 3 and 4). The Virtual Best Solver (VBS),
which considers the best algorithm for each instance separately, obtains much better results,
showing us that the algorithms have complementary performance. The difference between VBS and single
best (column $\delta$ of Table \ref{expTimeTable}) is more particularly important for rather small CPU
time limits. In particular, VBS solves 693 more instances than the single best (Glasgow 1) when the
limit is 100 ms. In many applications, it is important to have the fastest possible algorithm. For
example, in pattern recognition applications, we often have to solve subgraph isomorphism problems
for a very large number of graphs (in order to find a pattern image in a large database of target
images, for example), so that having an algorithm that is able to solve an instance in 100 ms
instead of 1000 ms makes a big difference. Therefore, it is important to select the best algorithm
for each instance.
